---
title: "SF1 Biomass"
author: "Stephen Formel"
date: "11/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Last updated: Nov 27, 2020
```

## Plant Biomass for SF1 project

Three quantities were measured.

1. AG = Aboveground biomass inside the experimental space
1. IR = Root biomass that was inside the experimental space
1. OT = All tissue (Above and below) that was outside the experimental space.

When we harvested the plants, it seemed like the ones with the most outside tissue were the ones with the most cracked jars, which makes sense.

## load libraries

```{r, message = FALSE}

library(readxl)
library(tidyverse)
library(ggplot2)
library(here)
library(brms)
library(modelr)
library(tidybayes)

```


## load and clean data

```{r}

df <- read_excel(here("data", 
                      "plant", 
                      "SF1_GC_data.xlsx"),
                 sheet = "biomass")

#Make total mass
df <- df %>%
  group_by(plantID) %>%
  mutate(total_mass = sum(dry_mass))


```


## Explore data

```{r}

#All tissue types
df %>%
  ggplot(aes(x = oil, 
             y = dry_mass, 
             fill = Tissue)) +
  geom_boxplot() +
  facet_grid(~ ifelse(soil=="L", "Live Soil", "Sterile Soil")) +
  xlab("Oil Added")

#Total Biomass
df %>%
  ggplot(aes(x = oil, 
             y = total_mass, 
             fill = soil)) +
  geom_boxplot() +
  xlab("Oil Added")

```

Oil might have affected plants growing in autoclaved soil, but it did not effect those growing in live soil.

```{r}

m <- lm(total_mass ~ soil*oil, data = df)
summary(m)
hist(resid(m))
shapiro.test(resid(m))

```

See if it can be transformed

```{r}

m <- lm(sqrt(total_mass) ~ soil*oil, data = df)
summary(m)
hist(resid(m))
shapiro.test(resid(m))

m <- lm(log(total_mass) ~ soil*oil, data = df)
summary(m)
hist(resid(m))
shapiro.test(resid(m))

```


Can't be transformed. Let's see if we can fit it with Bayes.

## Bayesian LM

Make model

```{r, message = FALSE, results = 'hide'}

b0 <- brm(total_mass ~ 1 + soil*oil, 
          data = df,
          family = "skew_normal",
          chains = 4, cores = 4,
          seed = 1)
```

Check model Results

```{r}

#https://bayesed-and-confused.netlify.app/post/model-fit-checks/

M <- b0
pp_check(object = M, type = "dens_overlay", nsamples = 100)
pp_check(M, type = "stat", stat = 'median', nsamples = 100)
pp_check(M, type = "stat", stat = 'mean', nsamples = 100)
pp_check(M,type = 'intervals', nsamples = 100)

plot(M)
as.data.frame(posterior_summary(M))

#R2
as.data.frame(bayes_R2(object = M,resp = NULL,summary = TRUE,robust = FALSE,probs = c(0.025, 0.975)))

#loo_R2
loo_R2(object = M)

```

Not a perfect model, but it does look like sterile soil increased the biomass, while oil decreased it.  Probably not interactive effect, however.

## Plot with tidy_bayes-----

```{r}

grid = df %>%
  data_grid(soil, oil)

fits = grid %>%
  add_fitted_draws(M)

preds = grid %>%
  add_predicted_draws(M)

df %>%
  ggplot(aes(x = total_mass, y = interaction(oil, soil))) +
  stat_halfeye(aes(x = .value), scale = 0.6, position = position_nudge(y = 0.175), data = fits) +
  stat_interval(aes(x = .prediction), data = preds) +
  geom_point(data = df) +
  scale_color_brewer() +
  theme_bw()

```

This shows posteriors of the mean of each group (black intervals and the density plots) and posterior predictive intervals (blue).

