---
title: "SF1 Plant Growth Rate"
author: "Stephen Formel"
date: "11/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Fitting non-linear models to growth rate.  Based on [Paine et al. (2012)](https://besjournals.onlinelibrary.wiley.com/doi/abs/10.1111/j.2041-210X.2011.00155.x)

and

http://www.statsathome.com/2017/06/07/fitting-non-linear-groth-curves-in-r/

## Load Libraries

```{r}

library(tidyverse)
library(readxl)
library(here)
library(brms)

set.seed(4)

```

Load and clean data
```{r}

df <- read_excel(here("data", 
                      "plant", 
                      "SF1_GC_data.xlsx"),
                 sheet = "morphology", na = "NA")

#convert to factor

df <- df %>% 
  mutate_if(is.character,as.factor)

#make biomass proxy, stem_ht (convert to mm) * stem diam

df <- df %>% 
  group_by(soil, oil, plantID, week, stem_number) %>% 
  summarise(biomass_proxy = stem_ht*10*stem_diam) %>%
  na.omit() %>% 
  group_by(soil, oil, plantID, week) %>% 
  summarise(sum_biomass_proxy = sum(biomass_proxy))


```

## Make Gompertz function

```{r}

fit.gompertz <- function(data, time){
  d <- data.frame(y=data, t=time)
  
  # Must have at least 3 datapoints at different times
  if (length(unique(d$t)) < 3) stop("too few data points to fit curve")
  
  # Pick starting values ###
  i <- which.max(diff(d$y))
  starting.values <- c(a=max(d$y), 
                       mu=max(diff(d$y))/(d[i+1,"t"]-d[i, "t"]), 
                       lambda=i)
  print("Starting Values for Optimization: ")
  print(starting.values)
  ##########################
  
  formula.gompertz <- "y~a*exp(-exp(mu*exp(1)/a*(lambda-t)+1))"
  nls(formula.gompertz, d, starting.values)
}

```


## Fit the Gompertz model

```{r}

(fit <- safe.fit.gompertz(df$week, df$biomass_proxy))

#singular fit, which means that starting values for estimation are bad.

plot(df$sum_biomass_proxy ~ df$week)

```




```{r}

###Log fit - be sure to use quotes around the variable names in the call
log.fit <- function(dep, ind, yourdata){
#Self-starting...

y <- yourdata[, dep]
x <- yourdata[, ind]

log.ss <- nls(y ~ SSlogis(x, phi1, phi2, phi3))

#C
C <- summary(log.ss)$coef[1]
#a
A <- exp((summary(log.ss)$coef[2]) * (1/summary(log.ss)$coef[3]))
#k
K <- (1 / summary(log.ss)$coef[3])

plot(y ~ x, main = "Logistic Function", xlab=ind, ylab=dep)
lines(0:max(x), predict(log.ss, data.frame(x=0:max(x))), col="red")

r1 <- sum((x - mean(x))^2)
r2 <- sum(residuals(log.ss)^2)

r_sq <- (r1 - r2) / r1

out <- data.frame(cbind(c(C=C, a=A, k=K, R.value=sqrt(r_sq))))
names(out)[1] <- "Logistic Curve"

return(out)
}

```


#Fit logistic curve

Looks like Gompertz isn't a good choice and logistic is probably better.  Good place to start:

https://stats.stackexchange.com/questions/47802/whats-the-most-pain-free-way-to-fit-logistic-growth-curves-in-r

https://bscheng.com/2014/05/07/modeling-logistic-growth-data-in-r/#:~:text=A%20logistic%20growth%20model%20can,for%20non%2Dlinear%20least%20squares.&text=This%20is%20because%20R%20will,needs%20a%20place%20to%20start.

```{r}

bp.mod <- nls(sum_biomass_proxy ~ phi1/(1+exp(-(phi2+phi3*week))),
 start=list(phi1 = 10000, phi2 = -1.096, phi3 = .002),data = df, trace=TRUE)


#need to make tibble into data frame for function to be happy
df.df <- as.data.frame(df)

(fit <- log.fit(dep = "sum_biomass_proxy", ind = "week", yourdata = df.df))

```

Gosh, non-linear models are a whole new can of worms.  Let's start more basically by looking at a GLM 

## Bayesian LM

Make model

```{r, message = FALSE, results = 'hide'}

b0 <- brm(sum_biomass_proxy ~ 1 + unlist(week) + soil*oil + (1|plantID), 
          data = df,
          family = "skew_normal",
          chains = 4, cores = 4,
          seed = 1)
```

Check model Results

```{r}

#https://bayesed-and-confused.netlify.app/post/model-fit-checks/

M <- b0
pp_check(object = M, type = "dens_overlay", nsamples = 100)
pp_check(M, type = "stat", stat = 'median', nsamples = 100)
pp_check(M, type = "stat", stat = 'mean', nsamples = 100)
pp_check(M,type = 'intervals', nsamples = 100)

plot(M)
as.data.frame(posterior_summary(M))

#R2
as.data.frame(bayes_R2(object = M,resp = NULL,summary = TRUE,robust = FALSE,probs = c(0.025, 0.975)))

#loo_R2
loo_R2(object = M)

```

```