---
title: "SF1 Soil Chemistry"
author: "Stephen Formel"
date: "11/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#last updated Oct 28, 2020 by Steve Formel

```

## Description

Basic soil chemistry was analyzed by LSU AgCenter.  This is to test whether there was a difference among the treatments at the beginning of the experiment and how conditions changed over the course of the experiment.

```{r, message = FALSE}

#import data and load libraries-----
library(ggfortify)
library(readxl)
library(plyr)
library(vegan)
library(tidyverse)
library(compositions)
library(ggplot2)
library(here)
library(cowplot)
```

Load and clean data

```{r}

df <- read_excel(path = here::here("data",
                             "soil_chem",
                             "SF1_soil_chem_all_data.xls"),
                 sheet = "data_cleaned")

#clean data----

#rename chem titles
colnames(df) <- c("sampleID", "soil", "oil", "time", "Calcium", "Copper", "Magnesium", "pH", "Phosphorus", "Potassium",  "Sodium", "Sulfur", "Zinc", "Percent Carbon", "Percent Nitrogen")

df$soil <- plyr:::revalue(df$soil, c("L"="Live", "S"="Autoclaved"))
df$oil <- plyr:::revalue(df$oil, c("N"="No Oil", "Y"="Oil Added"))

df$time <- factor(df$time, levels = c("pre-exp", "post-exp"))
df$Treatment <- interaction(df$soil, df$oil)

SF1_aes <- aes(shape = soil,
      fill = oil)

my.scale_aes <- list(scale_shape_manual(values = c(21,24)),
    scale_fill_manual(values = c("white", "black")),
    theme_bw(),
    guides(fill = guide_legend(override.aes=list(shape=21, size = 4)),
           shape = guide_legend(override.aes=list(size = 4))),
    labs(shape = "Soil",
         fill = "Oil"))

```


## Pre-experiment

Soil state at outset of experiment.

```{r}

df.pre <- subset(df, df$time=="pre-exp")

#PCA and plot pre-experiment only as biplot
#from https://towardsdatascience.com/principal-component-analysis-pca-101-using-r-361f4c53a9ff

df.PCA <- prcomp(df.pre[c(5:15)], center = TRUE, scale = TRUE)

p <- autoplot(df.PCA, 
         data = df.pre,
         loadings = TRUE, 
         loadings.colour = 'black',
         loadings.label.colour = 'black',
         loadings.label = TRUE,
         loadings.label.vjust = -1.5,
         loadings.label.hjust = -0.05,
         loadings.label.size = 3)

p$layers <- p$layers[-1]

pre <- p+ geom_point(mapping = SF1_aes) +
  my.scale_aes +
  geom_abline(intercept = -0.1, slope = -0.7, color="gray", 
              linetype="dashed", size=0.5)

#print
pre

```


## Post-Experiment

Soil state at experiment harvest.

```{r}

df.post <- subset(df, df$time=="post-exp")

#PCA and plot post-experiment only as biplot
#from https://towardsdatascience.com/principal-component-analysis-pca-101-using-r-361f4c53a9ff

df.PCA <- prcomp(df.post[c(5:15)], center = TRUE, scale = TRUE)


p <- autoplot(df.PCA, 
         data = df.post,
         loadings = TRUE, 
         loadings.colour = 'black',
         loadings.label.colour = 'black',
         loadings.label = TRUE,
         loadings.label.vjust = -1.5,
         loadings.label.hjust = -0.05,
         loadings.label.size = 3)
  
p$layers <- p$layers[-1] 
  
post <- p + geom_point(mapping = SF1_aes) +
  my.scale_aes +
  geom_abline(intercept = -0.1, slope = -0.6, color="gray", 
              linetype="dashed", size=0.5) +
  geom_abline(intercept = 0.1, slope = 1.5, color="gray", 
              linetype="dashed", size=0.5)

#print
post

```


## Plot Together

Plot pre- and post- experiment chemistry together

```{r}

top <- plot_grid(pre + theme(legend.position = "none"), post + theme(legend.position = "none"), ncol = 2, labels = "AUTO")

plot_grid(top, get_legend(post + theme(legend.position = "bottom")), ncol = 1, rel_heights = c(1,0.2))

```

## Relative abundance of chemistry

```{r, out.width = "100%"}

#CLR Transformations

#This transforms each number into it's relative abundance in that sample
comps <- acomp(df[,c(5:7,9:13)])

RA.df <- cbind(comps, df)

RA.df.gathered <- RA.df %>%
  gather(key = "Element", value = "Proportion", c(1:8))

RA.df.gathered %>%
  ggplot(aes(x = Element,
             y = Proportion,
             color = Element)) +
  geom_boxplot() +
  facet_grid(rows = vars(time),
             cols = vars(interaction(soil,oil))) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
```

But as I understand it, it's inappropriate to do the stats on this because it's closed.  Need to translate to CLR.

### Multivariate Methods

The central idea of the package following the coordinate approach of [Pawlowsky-Glahn(2003)] and [Pawlowsky-Glahn and Mateu-Figueras(2005)] is to transform the data by one of transforms into a classical multivariate dataset, to apply classical multivariate statistics and to back transform or interpreted the results afterwards in the original space.

According to https://www.rdocumentation.org/packages/rgr/versions/1.1.15/topics/clr all units should be the same, so the percents should be converted to ppm (or whatever is used).  

*pH won't be CLR transformed because it's not compositional relative to the other elements (it covaries, but I don't think that's considered compositional).*

However, this paper (10.1007/s12665-019-8248-6) just used clr transformed pH , but its not clear what they used as the geometric mean.  

## Aitchinson Distance of Chemistry

Re-analyze using Aitchison distance.

```{r}

#Convert %C and N to ppm
df$`Percent Carbon` <- df$`Percent Carbon`*10000
df$`Percent Nitrogen` <- df$`Percent Nitrogen`*10000

clr.df <- clr(df[,c(5:7,9:15)])

clr.df <- as.data.frame(clr.df)
clr.df$pH <- df$pH

pc <- princomp(x = clr.df)

pc$Loadings # The loadings as compositional vector
pc$loadings # The loadings in clr-space
df.pca <- pc$scores

aitchison.pca.plot <- cbind(na.omit(df), df.pca) %>%
  ggplot(aes(x = Comp.1,
             y = Comp.2)) +
  geom_point(mapping = SF1_aes) +
    facet_wrap(~ time) +
    my.scale_aes

p <- autoplot(pc, 
         data = df,
         loadings = TRUE, 
         loadings.colour = 'black',
         loadings.label.colour = 'black',
         loadings.label = TRUE,
         loadings.label.vjust = -1.5,
         loadings.label.hjust = -0.05,
         loadings.label.size = 3)
  
p$layers <- p$layers[-1] 

#set aspect ratio for plotting labels

p + geom_point(mapping = SF1_aes) +
  my.scale_aes +
  geom_abline(intercept = -0.075, slope = -0.6, color="gray", 
              linetype="dashed", size=0.5) +
  geom_abline(intercept = 0.1, slope = 1.5, color="gray", 
              linetype="dashed", size=0.5) +
  coord_fixed() +
  geom_text(data=data.frame(x = 0.15,y = 0.36), aes(x, y), label="Oiled", hjust = 0.5, vjust=0.5, angle = 58) +
  geom_text(data=data.frame(x = 0.19,y = 0.35), aes(x, y), label="Not Oiled", hjust = 0.5, vjust=0.5, angle = 58) +
  geom_text(data=data.frame(x = 0.32,y = -0.295), aes(x, y), label="Pre-Experiment", hjust = 0.5, vjust=0.5, angle = -30) +
  geom_text(data=data.frame(x = 0.35,y = -0.265), aes(x, y), label="Post-Experiment", hjust = 0.5, vjust=0.5, angle = -30)
```

## PERMANOVA on Aitchinson distance

```{r}

df.clr <- cbind(df[,c(1:4)], clr.df)

set.seed(1)
adonis(formula = clr.df ~ soil*oil*time, data = df.clr, permutations = 9999, method = "euclidean")

```

#Boxplot of CLR-transformed composition

```{r}

elements <- df.clr %>%
  gather(key = "Element", value = "clr_val", Calcium:'Percent Nitrogen') %>%
  ggplot(aes(x = fct_reorder(Element, clr_val, .fun='median'),
         y = clr_val)) +
  geom_point(mapping = SF1_aes,
              size = 3,
             position = position_dodge(width = 0.5)) +
  my.scale_aes +
  theme_bw() +
  labs(y = "Centered Log-Ratio",
       x = NULL) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~ time)

pH <- df.clr %>%
  ggplot(aes(x = time,
         y = pH)) +
  geom_point(mapping = SF1_aes,
              size = 3,
             position = position_dodge(width = 0.5)) +
  my.scale_aes +
  theme_bw() +
  labs(y = "pH",
       x = NULL)

```

## Plot What is intended to be figure S1

```{r, out.width = "100%"}

top <- plot_grid(elements + theme(legend.position = "None"), pH + theme(legend.position = "None"), rel_widths = c(1,0.5), labels = "AUTO")

plot_grid(top, get_legend(pH + theme(legend.position = "bottom")), nrow = 2, rel_heights = c(1,0.1))

```

