---
title: "SF1 Germination"
author: "Stephen Formel"
date: "11/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Last updated: Nov 27, 2020

```

Load Libraries

```{r, message=FALSE}

library(readxl)
library(here)
library(tidyverse)
library(ggplot2)
library(knitr)

```

Load and clean data

```{r}

df <- read_excel(here("data", 
                      "plant", 
                      "SF1_GC_data.xlsx"),
                 sheet = "germination")

```

Percent Germinated
```{r}

#Percent germinated at experiment end

df.mean <- df %>%
  filter(day_number==28) %>%
  group_by(soil, oil) %>% 
  summarise(perc_germ = (12-not_germinated) / 12) %>% 
  group_by(soil, oil) %>%
  summarise(mean_perc_germ = mean(perc_germ),
            sd = sd(perc_germ),
            se = sd(perc_germ)/sqrt(n()))

kable(df.mean)

df.mean %>%
  ggplot(aes(x = soil,
           y = mean_perc_germ,
           fill = oil)) +
  geom_col() +
  geom_errorbar(aes(ymin = mean_perc_germ-2*se,
                    ymax = mean_perc_germ+2*se),
                width = 0.2) +
  facet_wrap(~ oil)


```

## Stats on percent germinated

The idea here is to get a percentage of seeds germinated from each pot and then treat those as continuous variables.  I think this is ok as long as the variance is equal.  See [this discussion](https://stats.stackexchange.com/questions/11296/using-anova-on-percentages)

```{r}

df.perc <- df %>%
  filter(day_number==28) %>%
  group_by(soil, oil) %>% 
  summarise(perc_germ = (12-not_germinated) / 12) 

M <- lm(perc_germ ~ soil*oil, data = df.perc)

summary(M)
hist(resid(M))
shapiro.test(resid(M))

```

Check heteroscedasticity and normality

```{r}

df %>%
  filter(day_number==28) %>%
  ggplot(aes(x = soil,
           y = 12-not_germinated,
           color = oil)) +
  geom_boxplot()


```

Looks reasonable to me.

## Germination Rate

```{r}

M <- glm(perc_germ ~ soil*oil, family = quasibinomial, 
         weights = rep(12, nrow(df.perc)),
          data = df.perc)

#Adding weights doesn't seem to make a diference in results.

summary(M)
plot(M$residuals, M$fitted.values)


```

Whether soil was autoclaved or not is barely significant. I have a feeling if I remove the two replicates that didn't germinate any seeds, this will disappear.

```{r}

df.perc.no.zero <- df.perc %>% 
  filter(perc_germ!=0)

M <- glm(perc_germ ~ soil*oil, family = quasibinomial,
          data = df.perc.no.zero)

summary(M)

```

It only brings the p-value up to 0.08, so perhaps there really is something there.  I'll have to think on that.

## Plot of LOESS smoothing on germination

Nice [explanation](https://www.statisticshowto.com/lowess-smoothing/) of how LOESS smoothing works.

```{r}

ggplot(data = df,
       aes(x = day_number,
           y = 12 - not_germinated,
           shape = factor(oil),
           color = factor(soil))) +
  geom_jitter() +
  geom_smooth() +
  theme_bw()

```

